---
title: "Report on Rating Predictions for Movielens"
author: "Oliver Brueckner"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a report on rating predictions for the movielense data set. After some data exploration we will follow three approaches to reach our goal: Make predictions for the ratings in the validation set with an RMSE <= 0.8649.

## Inspecting the Data

The data was loaded with the code given in the course. A first view on the data shows that there is a huge dataset with more than 9 million ratings on movies. These ratings come from 69,878 users and are given to 10,677 movies. These numbers show, that not every movie is rated by every user. 

```{r echo=FALSE}
load("mydata.rda")
```

```{r loading-libs, message=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
```


```{r echo=FALSE}
edx_meta
```

If we have a closer look to the distribution of the Ratings we see, that they tend to be .0 instead of .5 The average rating is about 3.51 with a standard deviation of 1.06.

```{r echo=FALSE}
# What is the Distribution of Ratings?
rating_count%>%
  ggplot(aes(rating, n)) + 
  geom_point()+
  scale_y_log10() +
  ggtitle("Ratings")
```






You can also embed plots, for example:

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
